A paper is to write

### 2011-02-25 workflow.Rnw
Next

\title{Collaboration}

\author{Jake Bowers\thanks{Many thanks to Kevin Quinn for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.}}

@@

An idea is born in a coffee shop, a seminar, a quiet walk. It is winter 2011. This idea inspires a seminar paper in the spring. And a conference paper arises from the seminar paper in 2012. A dissertation chapter descends from the conference paper in 2013. Other dissertation chapters take up 2014. A submission to a journal happens in 2015. Revision and resubmission wait until 2017 while harried editors, reviewers and author strive to balance teaching, service, and life. In 2018 a child is born and a paper is published. In 2020 a graduate student in a coffee shop reads the paper and has an idea that challenges the analysis in the paper. What would happen if only the author had thought to control for X? Or run one more MCMC chain? Or chosen a different likelihood function? Will the United Nations (now eager to act based the paper) make a wrong move? Time to go back to the analyses run sometime between 2011 and 2020 to check. Easy. Right? Collaboration between the past self and the present self should be simple --- after all you are the same person in 2011 and 2020, right? And the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right? And you know exactly which bit of code produced which table and which figure, right? And you wrote your code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?

Just in case one has some doubt about answering these questions, this piece aims to amplify some of what we already ought to know \citet{nagler1995coding}, and to update some of those ideas given current practices, platforms, and propensities.

\vspace{4em}

@@

Here is some advice that is not currently emphasized but which might be profitably added to previous advice.

@@

An even nicer way to ensure that figures and tables are reproducible or easily modifiable by your older and wiser but more harried self is to use literate programming. For example, one may insert the following chunk of code into either a LaTeX or an OpenOffice document, ask R to process the document, and then produce a pdf file in which the code has disappeared to be replaced with an actual figure:

<<read.data>>=

@@

\section{Fishing rods and not fish}
The source code of any pdf document exchanged by the group must be available and executable. Say you run some command and then make a table or a graph. Then you produce a nice little report on your work for use among the working group. Eventually we want to use pieces of that report (tables, graphs, paragraphs) in a publishable paper. Thus, that report must not have been created using copying and pasting. It must be created using LaTeX (or HTML or OpenOffice or something else that the working group standardizes on). And, in fact, it must be created using the reproducible markup system like Sweave (if you are using LaTeX and R as your working group tools). Alternatively, you can put comments into your report detailing where the figures and tables came from (i.e. which command file produced them). Or include details about which code produced which figures in your MANIFEST.txt.

In order to know that ``that plot on page 10'' is from a file called ``fig1.pdf'' you need to document it somehow. I can imagine a system with very disciplined use of Word/OpenOffice --- perhaps using special auxiliary ``listoffigures.txt'' files or something. In my experience, it is much easier to just use LaTeX (which can be used in a very user friendly way via LyX).

@@

Group work requires version control.\footnote{See X for more
  discussion of what version control is.} Many people are familiar
with ``Track Changes'' feature in OpenOffice/Word/Pages or the fact that
Dropbox allows one to recover previous versions of files. These are
both kinds of version control. When collaborating with yourself or
others, it is useful to see what has changed, to feel free to
experiment and then to dump the experiment in favor of previous work,
to have multiple ``releases'' of the same document (one to MPSA, one
to APSR, one to your parents) without requiring that spawn many copies
of the same document and risk confusion and clobbering. Clobbering is
what happens when your future self or your other collaborators saves
an old version of a file over a new version --- erasing good work by
accident.

@@

Lots of people are thinking about ``reproducible research'' and ``literate programming'' these days. Google those terms. I have experimented with three systems so far: (1) for one paper I simply included my Sweave document and data files into a compressed archive; (2) for another more computing intensive paper...

[examples: Examples of reprodible research using R vignette system [jake's Manifest Effects paper?], and a more complex system that required more time instensive simulation using a Makefile [jake and ben and marks's JASA archive]?]

@@

The time-frames over which collaboration are required --- whether among a group of people working together or within a single scholar's productive life --- are much longer than any given version of any given software will easily exist. All except for plain text. Thus, even as we extol version control systems, one must have a way to ensure future access to them in a form that will still be around when the cockroaches take over political science departments (and thereby finally make C-x M-x C-c  a reasonable way to instruct Emacs to do something [insert url citing this argument --- that Emacs was really made for cockroaches or other six legged creatures.]

In the end, following these practices allows your analyses of your collections to be credible --- and, in the case of criticism, you can always tell antagonists to go redo your analysis if they really don't believe you.

"[I]f the empiricalbasis for an articleor book cannotbe reproduced,of what use to the discipline are its conclusions? What purposedoes an articlelike this serve? \cite[445]{king1995replication}

@@

\end{document}

>>>>

\title{A paper is to write}

\author{Jake Bowers\thanks{Many thanks to Kevin Quinn and Cara Wong for direct help on this document, Mika LaVaque-Manty, Ben Hansen, and Mark Fredrickson for many useful discussions on this topic, and Kieran Healy for inspiration from afar.}}

@@

An idea is born in a coffee shop, a seminar, a quiet walk. It is a cold winter in 2011, but the idea energizes the student. The idea is why he went to graduate school.

This idea inspires a seminar paper in the spring. And a conference paper arises from the seminar paper in collaboration with another student in 2012. A dissertation chapter descends from the conference paper in 2013. Other dissertation chapters take up 2014. A submission to a journal with the original co-author and a new collaborator happens in 2015. Revision and resubmission wait until 2017 while harried editors, reviewers and authors strive to balance teaching, service, and life. The three collaborators have luckily found work in three different universities around the world. In 2018 a child is born and a paper is published. In 2020 a graduate student in a coffee shop reads the paper and has an idea that challenges the analysis in the paper. What would happen if only the first author had thought to control for X? Or run one more MCMC chain? Or chosen a different likelihood function? Will the United Nations (now eager to act based the paper) make a wrong move?

The first author convenes a three way video conference with the other collaborators while commuting home in his flying car. The group must go back to the analyses. Which ones? The ones from 2011? Or 2012? Or 2018? Where are the files? The next day, one member of the group finds some of the files. Now the checks and re-analyses should be easy. Right? The student-now-faculty member is the same person and should remember the reason for those bits of code (or at least should remember which series of mouse clicks were used to produce the numbers for that crucial table). And, of course, the way Microsoft Word/Stata/SPSS/R/LISREL stores files and the way that your Windows/Mac machine reads and writes them is the same, right? (And Windows and Mac are still relevant operating systems.) And the group knows exactly which bit of code produced which table and which figure, right? And they wrote their code following Nagler's Maxims \citet{nagler1995coding} and being following King's Replication Standard \citet{king1995replication} right?

Just in case one has some doubt about answering these questions, this piece aims to amplify some of what we already ought to know from \citet{nagler1995coding} and \citet{king1995replication}, and to update some of those ideas given current practices, platforms, and propensities.

\section{Why should we believe each other?}

@@

@@

An even nicer way to ensure that figures and tables are reproducible or easily modifiable by your older and wiser but more harried self is to use literate programming. For example, to produce the Figure~\ref{fig:giniprot}, I insert the following chunk of code into the  LaTeX document, and then I asked R to process the document, and then produce a pdf file in which the code has disappeared to be replaced with an actual figure.\footnote{One can insert the same code chunk in an OpenOffice document and use the odfWeave package to produce OpenOffice documents in which figures and tables replace code. One may also use the LyX interface to LaTeX to write LaTeX documents within a WYSIWYG environment.}

<<read.data,echo=FALSE,results=hide>>=

@@

\section{Fishing rods and not fish}
The source code of any pdf document exchanged by the group must be available and executable. Say you run some command and then make a table or a graph. Then you produce a nice little report on your work for use among the working group. Eventually we want to use pieces of that report (tables, graphs, paragraphs) in a publishable paper. Thus, that report must not have been created using copying and pasting. It must be created using LaTeX (or HTML or OpenOffice or something else that the working group standardizes on). And, in fact, it must be created using the reproducible markup system like Sweave (if you are using LaTeX and R as your working group tools) \citep{Leis:2005}. Alternatively, you can put comments into your report detailing where the figures and tables came from (i.e. which command file produced them). Or include details about which code produced which figures in your MANIFEST.txt.

In order to know that ``that plot on page 10'' is from a file called ``fig1.pdf'' you need to document it somehow. I can imagine a system with very disciplined use of a word processor (like Open Office or Word) --- perhaps using special auxiliary ``listoffigures.txt'' files or something. In my experience, it is much easier to just use LaTeX (which can be used in a very user friendly way via LyX).

@@

Group work requires version control.\footnote{See X for more
  discussion of what version control is.} Many people are familiar
with ``Track Changes'' feature in modern WYSIWYG word processors or the fact that
Dropbox allows one to recover previous versions of files. These are
both kinds of version control. When collaborating with yourself or
others, it is useful to see what has changed, to feel free to
experiment and then to dump the experiment in favor of previous work,
to have multiple ``releases'' of the same document (one to MPSA, one
to APSR, one to your parents) without requiring that spawn many copies
of the same document and risk confusion and clobbering. Clobbering is
what happens when your future self or your other collaborators saves
an old version of a file over a new version --- erasing good work by
accident.

@@

Lots of people are thinking about ``reproducible research'' and ``literate programming'' these days. Google those terms. Of course the devil is the details: Here I list a few of my own attempts at enabling reproducible research. You'll find many other inspiring examples on the web.

I have experimented with three systems so far: (1) for one paper we simply included a Sweave document and data files into a compressed archive \cite{bowers2005dataverse}; (2) for another more computing intensive paper we assembled a set of files which enabled reproduction of our results using the ``make'' system \citep{bowers2008dataverse}; and (3) recently I have tried the ``compendium'' approach \citep{gentleman2005reproducible,gentleman2007statistical} which embeds an academic paper with the R package system \cite{bowers2011dataverse}. The benefit of this last system is that one is not required to have access to unix: one may download, run, and view the paper from within R --- which is platform independent.

@@

The time-frames over which collaboration are required --- whether
among a group of people working together or within a single scholar's
productive life --- are much longer than any given version of any
given software will easily exist. All except for plain text. Thus,
even as we extol version control systems, one must have a way to
ensure future access to them in a form that will still be around when
the cockroaches take over political science departments.\footnote{
  (and thereby finally make Ctrl-a Ctrl-x Esc-x Ctrl-c a reasonable
  way to instruct Emacs to do something
  (\url{http://kieran.healy.usesthis.com/} )}

In the end, following these practices allows your analyses of your collections to be credible --- and, in the case of criticism, you can always tell antagonists to go redo your analysis if they really don't believe you.  Let me end with the most pithy statement in favor of the idea that collaboration is at the heart of our endeavor that I have found:

"[I]f the empirical basis for an article or book cannot be reproduced, of what use to the discipline are its conclusions? What purpose does an article like this serve?'' \cite[445]{king1995replication}

@@

\end{document}
<<>>=
options(prompt="> ",continue="+ ")
@
